\frontmatter 

% title page, list of tables, list of figures
\input{frontmatter_title}

%Brief overview of the document, acknowledgement, disclaimer
\chapter{Preface}

\section*{Who Should Read This Lab Projects Manual?}
This lab projects manual is written for students who are taking Electrical and Computer Engineering (ECE) Systems Programming and Concurrency course ECE 252 in the University of Waterloo.

\section*{What is in This Lab Projects Manual?}

The first purpose of this document is to provide the descriptions of each laboratory project
\footnote{We use lab project, lab, project interchangeably in this document}.
The second purpose of this document is a quick reference guide of the relevant development tools 
for completing laboratory projects. This manual is divided into three parts. 

Part I describes the lab administration policies 

Part II is a set of course laboratory projects as follows.

\begin{itemize}
    \item Lab1: Introduction to systems programming in Linux computing environment
    \item Lab2: Multi-threaded concurrency programming with blocking I/O
    \item Lab3: Inter-process communication and concurrency control
    \item Lab4: Parallel web crawling
    \item Lab5: Single-threaded concurrency programming with asynchronous I/O  
\end{itemize}

Part III is a quick reference guide of the Linux software development tools. We will be using Ubuntu operating system. Materials in this part needs to be self-studied before lab starts.
The main topics are as follows.

\begin{itemize}
    \item Linux hardware environment
    \item Editors
    \item Compiler 
    \item Debugger 
    \item Utility to automate build
    \item Utility for version control
\end{itemize}


% A C K N O W L E D G E M E N T S
% -------------------------------

\section*{Acknowledgments}

We are grateful that Professor Patrick Lam shared his ECE 459  projects with us. Eric praetzel has provided continuous IT support, which makes the Linux computing environment available to our students.

We would like to sincerely thank our students who took ECE 254 and ECE 459 courses in the past few years. They provided constructive feedback every term to make the manual more useful to address problems that students would encounter when working on each lab project. 

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main_book"
%%% End:
